---
title: "Building a Content Delivery Pipeline for a Website in R"
subtitle: "Embedding web components from R into a content management system"
author:
  - "Gregor de Cillia"
  - "Alexander Kowarik"
  - "Bernhard Meindl"
output: 
  slideSTAT::slideSTAT:
    h1_slides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

## Abstract

[Statistics Austria] is currently working towards the release of a new website.
An important part of it, namely **graphs** and **tables**, will be created using [R]

Content creators will create **graphs** and **tables** via a [shiny app] and transfer them to the content management system (**CMS**).
The transfer uses a [plumber] REST API under the hood

## Outline

\includegraphics{workflow.png}

Parts of the presentation

1. Showcase the **app** and **components**
2. Remarks about the **data import**
3. Mechanism to **transfer** components from the **app** to the **CMS**

# Components and App

## Components

The components that can be delivered by the pipeline are **graphs** and **tables**

Both the graphs and tables are **interactive** in the sense that they provide hover and/or click events

* Graphs are generated with [highcharts.js] and the corresponding R package [highcharter]
* Tables are created using [datatables.js] which provides features like **sorting**, **searching** and more

Interactive demos can be found in the [html version of these slides]

## Component Builder App

The component builder **app** is a [shiny] application which is hosted on [RStudio Connect].
It guides the content creator through the component generation process

- Data can be **imported** from different sources
- Components can be **created** and **previewed**.
  Sensible defaults make it possible do this with a few clicks
- An **export menu** makes the component accessible for the **CMS**

# Data Import

## Data Import

Currently, there are two datasources which are compatible the component builder **app**: [STATcube] and [Open Government Data]

Both data sources are imported using the open source package [STATcubeR] which defines a common **interface** for the data sources

```r
# online docs: statistikat.github.io/STATcubeR
remotes::install_github("statistikat/STATcubeR")
```

## Open Government Data

The [Open Data Portal] of Statistics Austria provides a series of datasets following the **open data** principles

About **300 datasets** are compatible with the component builder **app** which provides a coverage of about **95%**

OGD data can be imported from the **app** by choosing one of the datasets from a selection menu

More information can be found at https://statistikat.github.io/STATcubeR/articles/od_table.html

## STATcube

STATcube allows you to **generate** custom tables from databases and **export** them in different formats

One export option, the "[json API request]", is compatible with the **component builder app**.
The exported json file can be provided to the app via a [shiny fileInput]

The app then fetches data from the **STATcube REST API**

More information can be found at https://statistikat.github.io/STATcubeR/articles/sc_table.html

# Transfer Mechanism

## Transfer: App to CMS

\includegraphics{workflow_api.png}

In order to make the **graphs** and **tables** available for the **CMS**, a **REST API** is used.
The **app exports** the components to a **database** and the **CMS imports** the components

The API is built with [plumber] and hosted on [RSConnect]

## Export: App to Database

When a user **exports** a graph or table, the app performs a **/POST** request against the component API

The component is submitted as a [binary] object in the request body and added to a **BLOB** field in the **component database**

The database also captures **metadata** about the component such as the author and a timestamp

##

The **`/GET`** method of the **component API** lists all components in a json format. A tabular representation is provided below

```r
httr::GET("rsconnect.local/component-api")
```

```{r, echo=FALSE, out.width="90%"}
library(magrittr)
data.frame(
  Key = c("001", "002", "003", "004"),
  "Created at" = c("2021-09-19/11:43", "2021-09-19/14:17", "2021-09-20/06:34", "2021-09-21/13:29"),
  Creator = c("decill", "meindl", "decill", "kowa"),
  "Custom Label" = c("Cancer Types by Year", "Household Forecast", "Structure of Earnings", "Tourism: Correlations"),
  "Data Source" = c("OGD_krebs_ext_KREBS_1", "OGD_f1741_HH_Proj_1", "OGD_veste301_Veste301_1", "detouextregsai"),
  "Graph" = c("NULL", "Timeseries", "Barplot", "Scatterplot"),
  "Table" = c("175x3", "NULL", "122 x 7", "02 x 9")
) %>%
  knitr::kable() %>%
  kableExtra::kable_styling(font_size = 5.8)
```

Each row represents a **graph**, a **table** or both. Notice the **`NULL`** values in the last two rows

The columns **Creator** and **Data Source** reference the user and dataset behind a component

## Import: Database to CMS

Single items can be requested with `/GET table` and `/GET graph`.
They return a **rendered** version of a component for the **CMS**

The json payload of the response contains a `<div>` or `<table>` tag and a `<script>` tag which define the component

```r
# get the component from the api
response <- httr::GET(
  "rsconnect.local/component-api/graph?Key=001")
payload  <- httr::content(response)

# fill a html template
fill_table_template(
  # pass a javascript string
  script    = payload['script'],
  # pass container (<div>) as html string
  container = payload['container']
)
```

## Template example

The **template** defines where the component (graph) is inserted and loads all necessary **js/css dependencies**

```html
<head>
  <script src="highcharts.js"/>
  <script src="rsconnect.local/component-api/utils.js"/>
</head>
 
<body>
  <h1>Graph template</h1>
  <!-- define placeholder container for the graph -->
  ${container}
 
  <!-- fill container using a script tag -->
  ${script}
</body>
```

**utils.js** is a helper package that defines functions for **layouting** of components, **formatting tooltips**, etc.

# Conclusion

## Most Valuable Packages

Frameworks

* [shiny] defines the component builder **app**
* The component **API** is generated with [plumber]

Front End

* [highcharter] builds the **graphs**
* All **tables** are generated using [datatables.js]

Back End

* [httr] communicates with the **plumber API** and the **STATcube API**
* Data **import** is performed with the [STATcubeR] package

## Open Source Roadmap

Currently, the package [STATcubeR] is available on github.
It will be released to [CRAN] when the [STATcube] REST API becomes available for external users

The internal packages **STATgraph** and **STATtable**, which create graphs and tables based on the **STATcubeR** data interface, might become open-source once the pipeline is fully integrated in the **CMS** workflow

## Thank you for your attention

More information as well as the source code for these slides can be found at https://github.com/GregorDeCillia/pipedream

A video with this presentation is available at https://youtu.be/kjUtApjK6XE?t=215

[Statistics Austria]: https://www.statistik.at
[shiny]: https://shiny.rstudio.com/
[shiny app]: https://shiny.rstudio.com/
[R]: https://www.r-project.org/
[plumber]: https://www.rplumber.io/
[highcharts.js]: https://www.highcharts.comm/
[highcharter]: https://jkunst.com/highcharter/
[datatables.js]: https://datatables.net/
[RStudio Connect]: https://www.rstudio.com/products/connect/
[RSConnect]: https://www.rstudio.com/products/connect/
[STATcube]: https://www.statcube.at/
[Open Government Data]: https://data.statistik.gv.at
[Open Data Portal]: https://data.statistik.gv.at
[json API request]: https://docs.wingarc.com.au/superstar/9.10/open-data-api/use-superweb2-to-build-api-queries
[shiny fileInput]: https://shiny.rstudio.com/reference/shiny/1.6.0/fileInput.html
[binary]: https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/serialize
[httr]: https://httr.r-lib.org/
[CRAN]: https://cran.r-project.org/
[STATcubeR]: https://statistikat.github.io/STATcubeR/
[html version of these slides]: https://GregorDeCillia.github.io/pipedream/